Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-17 arxiv/proof_gaussian_case.tex
appendix

To study $W$'s resolvent, we observe that $W$ is equal to $\AF\AF^{\top}$ for a $p$ by $n_1 + n_2$ matrix
	\[ \AF := (n_1+ n_2)^{-1/2} [\Lambda U^\top (Z^{(1)})^\top,V^\top (Z^{(2)})^\top]. \]
%}\HZ{what does this trick mean? use less technical words},
%This idea dates back at least to Girko, see e.g., the works \cite{girko1975random,girko1985spectral} and references therein.
Consider the following symmetric block matrix whose dimension is $p + n_1 + n_2$
%which is a linear function of $(Z^{(1)})$ and $(Z^{(2)})$:
%\begin{definition}[Linearizing block matrix]\label{def_linearHG}%definiton of the Green function
%We define the $(n+N)\times (n+N)$ block matrix
 \begin{equation}\label{linearize_block}
    H \define \left( {\begin{array}{*{20}c}
   0 & \AF  \\
   \AF^{\top} & 0
%   {Z^{(2)}V} & 0 & 0
   \end{array}} \right).
 \end{equation}
For this block matrix, we define its resolvent as
$$G(z) \define \left[H - \begin{pmatrix}z\id_{p\times p}&0\\ 0 & \id_{(n_1+n_2)\times (n_1+n_2)} \end{pmatrix}\right]^{-1},$$
%as the resolvent of $H$,
for any complex value $z\in \mathbb C$.
Using Schur complement formula for the inverse of a block matrix, it is not hard to verify that
	\begin{equation*} %\label{green2}
	  G(z) =  \left( {\begin{array}{*{20}c}
			(W- z\id)^{-1} & (W - z\id)^{-1} \AF  \\
      \AF^\top (W - z\id)^{-1} & z(\AF^\top \AF - z\id)^{-1}
		\end{array}} \right).%\quad \cal G_R:=(W^\top W - z)^{-1} ,
  \end{equation*}

@@

\paragraph{Bias asymptotic limit.}
 For the bias limit in equation \eqref{lem_cov_derv_eq}, we show that it is governed by the derivative of $(W - z\id)^2$ with respect to $z$ at $z = 0$.
First, we can express the empirical bias term in equation \eqref{lem_cov_derv_eq} as %$W$
\begin{align}\label{calculate G'}
	(n_1 + n_2)^2 \hat{\Sigma}^{-1}\Sigma^{(2)}\hat{\Sigma}^{-1} = {\Sigma^{(2)}}^{-1/2} V W^{-2} V^{\top} {\Sigma^{(2)}}^{-1/2}.
\end{align}
Let $\cal G(z):=(W-z\id )^{-1}$ denote the resolvent of $W$.
Our key observation is that $\frac{\dd{\cal G(z)}}{\dd z} =  \cal G^2(z)$.
Hence, provided that the limit of $(W - z\id)^{-1}$ is $(a_1(z) \Lambda^2 + (a_2(z) - z) \id)^{-1}$ near $z = 0$, the limit of $\frac{\dd{\cal G(0)}}{\dd z}$ satisfies that
\begin{align}\label{cal G'0}
	\frac{\dd \cal G(0)}{\dd z} \to \frac{-\frac{\dd a_1(0)}{\dd z}\Lambda^2 - (\frac{\dd a_2(0)}{\dd z} - 1)\id}{(a_{1}(0)\Lambda^2 + a_{2}(0)\id_p)^2}.
\end{align}
To find the derivatives of $a_1(z)$ and $a_2(z)$, we take the derivatives on both sides of the system of equations \eqref{selfomega_a}.
%\begin{align*}
%	\frac{\dd a_1(z)}{\dd z} + \frac{\dd a_2(z)}{\dd z} = -\frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{1}{\lambda_i^2 a_1 + a_2},
%	\frac{\dd a_1(z)}{\dd z} + \frac{1}{n_1 + n_2}\sum_{i=1}^p \frac{\lambda_i^2 (a_1'(z) a_2 - a_2'(z) a_1)}{(\lambda_i^2 a_1 + a_2)^2} = -\frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 a_1}{(\lambda_i^2 a_1 + a_2)^2}
%\end{align*}
Let $a_3 = - \frac{\dd a_1(0)}{\dd z}$ and $a_4 = - \frac{\dd a_2(z)}{\dd z}$.
%then taking implicit differentiation of equation \eqref{selfomega_a}
One can verify that $a_3$ and $a_4$ satisfy the self-consistent equations in \eqref{eq_a34extra} (details omitted).
Applying equation \eqref{cal G'0} to equation \eqref{calculate G'}, we obtain the asymptotic limit of the bias term.
%\begin{align}
%& n^2\bignorm{\Sigma_2^{1/2} \bigbrace{ (X^{(1)})^{\top}(X^{(1)}) + (X^{(2)})^{\top}(X^{(2)}) }^{-1} \Sigma_1^{1/2} w}^2 \nonumber\\
%&\approx  w^\top \Sigma_1^{1/2}\Sigma_2^{-1/2}V\frac{a_3\Lambda^2 +(1+ a_4)\id_p}{(a_{1}\Lambda^2 + a_{2}\id)^2}V^\top \Sigma_2^{-1/2}\Sigma_1^{1/2}w= w^{\top} \Pi_\bias w, \label{calculatePibias}
%\end{align}
%where in the last step we used $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ and $V \Lambda^2 V^\top=M^\top M$. This concludes equation \eqref{lem_cov_derv_eq}.

@@

Next, we introduce the bounded support assumption for a random matrix.
We say that a random matrix $Z \in\real^{n \times p}$ satisfies the {\it{bounded support condition}} with $q$ or $Z$ has support $q$ if
\begin{equation}
	\max_{1\le i \le n, 1 \le j \le p}\vert Z_{i j} \vert \prec q. \label{eq_support}
\end{equation}
As shown in the example above, if the entries of $Z$ have finite moments up to any order, then $Z$ has bounded support $1$.
More generally, if the entries of $Z$ have finite $\varphi$-th moment, then using Markov's inequality and a simple union bound we get that %\HZ{I don't fully get the equation below}
\begin{align}
	\P\left(\max_{1\le i\le n, 1\le j \le p}|Z_{i  j}|\ge (\log n) n^{\frac{2}{\varphi}}\right) &\le \sum_{i=1}^n \sum_{j=1}^p \P\left(|Z_{i j}|\ge (\log n) n^{\frac{2}{\varphi}}\right)  \nonumber\\
	&\le \sum_{i=1}^n \sum_{j=1}^p \frac{C(\varphi)}{\left[(\log n) n^{\frac{2}{\varphi}}\right]^\varphi} = \OO((\log n)^{-\varphi}).\label{Ptrunc}
	\end{align}
In other words, $Z$ has bounded support $q=n^{\frac{2}{\varphi}}$. %with high probability.
%This leads to the exponent $\frac{\varphi-4}{2\varphi}$ in Fact \ref{lem_minv}. %This fact will be used in the proof of Corollary \ref{main_cor} below.
%The above discussion shows that the bounded support condition provides great flexibility in dealing with finite moment conditions.
%\begin{assumption}%\label{assm_big1}
%As in Section \todo{add reference}, we consider random matrices $X^{(1)}=Z^{(1)}\Sigma_1^{1/2}$ and $X^{(2)}=Z^{(2)}\Sigma_2^{1/2}$, where $Z^{(1)}$ and $Z^{(2)}$ are independent $n_1\times p$ and $n_2\times p$ random matrices with i.i.d. entries of zero mean and unit variance. Recall that $\rho_1= n_1/p$ and $\rho_2=n_2/p$. We assume that for a small constant $0<\tau<1$,
%\HZ{We have defined $\rho_1,\rho_2$ already in sec 2. Consider using ``Recall that $\rho_1 =, \rho_2 =$.''}
%\be\label{assm2}
%0\le \rho_1 \le \tau^{-1}, \quad 1+\tau \le \rho_{2} \le \tau^{-1}.
%\ee
%Recall that the singular value decomposition of $M = \Sig_1^{1/2} \Sig_2^{-1/2}$ is $U \Lambda V^{\top}$. We assume that
%\begin{equation}\label{assm32}
%\tau \le \lambda_p \le \lambda_1 \le \tau^{-1} .%, \quad \max\left\{\pi_A^{(n)}([0,\tau]), \pi_B^{(n)}([0,\tau])\right\} \le 1 - \tau .
%\end{equation}
%Finally, recalling the notations in equation \eqref{labelZ}, we assume that $Z^{(1)}$ and $Z^{(2)}$ satisfy
%\begin{equation}
%n^{-1/2}\max_{i\in \cal I_0,\mu \in \cal I_1}\vert Z^{(1)}_{\mu i}\vert \prec q,\quad n^{-1/2}\max_{i\in \cal I_0,\nu \in \cal I_2}\vert Z^{(2)}_{\nu i}\vert \prec q, \label{eq_support}
%\end{equation}
%for some deterministic parameter $q$ satisfying $ n^{-{1}/{2}} \leq q \leq n^{- \tau} $ for a small constant $\tau>0$.
%Here we recall the notations in \eqref{labelZ}.
 %and (\ref{assm2}). We assume that $T$ is an $M\times M$ deterministic diagonal matrix satisfying (\ref{simple_assumption}) and (\ref{assm3}).
%\end{assumption}
%\HZ{This is restating what's stated above again. Remove redundant statements (unless there's something critical missing in Sec 2).} [\cor FY: The setting in sec 1 is too vague for the purpose of this section in the sense that it is hard to find the exact properties we need. In this section, I want to state rigorous statements with a reference to an exact assumption. Moreover, the above assumption contains more.\nc]

>>>>

To study $W$'s resolvent, we observe that $W$ is equal to $\AF\AF^{\top}$ for a $p$ by $n_1 + n_2$ matrix
	\be\label{defn AF} \AF := (n_1+ n_2)^{-1/2} [\Lambda U^\top (Z^{(1)})^\top,V^\top (Z^{(2)})^\top]. \ee
%}\HZ{what does this trick mean? use less technical words},
%This idea dates back at least to Girko, see e.g., the works \cite{girko1975random,girko1985spectral} and references therein.
Consider the following symmetric block matrix whose dimension is $p + n_1 + n_2$
%which is a linear function of $(Z^{(1)})$ and $(Z^{(2)})$:
%\begin{definition}[Linearizing block matrix]\label{def_linearHG}%definiton of the Green function
%We define the $(n+N)\times (n+N)$ block matrix
 \begin{equation}\label{linearize_block}
    H \define \left( {\begin{array}{*{20}c}
   0 & \AF  \\
   \AF^{\top} & 0
%   {Z^{(2)}V} & 0 & 0
   \end{array}} \right).
 \end{equation}
For this block matrix, we define its resolvent as
$$G(z) \define \left[H - \begin{pmatrix}z\id_{p\times p}&0\\ 0 & \id_{(n_1+n_2)\times (n_1+n_2)} \end{pmatrix}\right]^{-1},$$
%as the resolvent of $H$,
for any complex value $z\in \mathbb C$.
Using Schur complement formula for the inverse of a block matrix, it is not hard to verify that
	\begin{equation} \label{green2}
	  G(z) =  \left( {\begin{array}{*{20}c}
			(W- z\id)^{-1} & (W - z\id)^{-1} \AF  \\
      \AF^\top (W - z\id)^{-1} & z(\AF^\top \AF - z\id)^{-1}
		\end{array}} \right).%\quad \cal G_R:=(W^\top W - z)^{-1} ,
  \end{equation}

@@

\paragraph{Bias asymptotic limit.}
 For the bias limit in equation \eqref{lem_cov_derv_eq}, we show that it is governed by the derivative of $(W - z\id)^2$ with respect to $z$ at $z = 0$.
First, we can express the empirical bias term in equation \eqref{lem_cov_derv_eq} as %$W$
\begin{align}\label{calculate G'}
	(n_1 + n_2)^2 \hat{\Sigma}^{-1}\Sigma^{(2)}\hat{\Sigma}^{-1} = {\Sigma^{(2)}}^{-1/2} V W^{-2} V^{\top} {\Sigma^{(2)}}^{-1/2}.
\end{align}
Let $\cal G(z):=(W-z\id )^{-1}$ denote the resolvent of $W$.
Our key observation is that $\frac{\dd{\cal G(z)}}{\dd z} =  \cal G^2(z)$.
Hence, provided that the limit of $(W - z\id)^{-1}$ is $(a_1(z) \Lambda^2 + (a_2(z) - z) \id)^{-1}$ near $z = 0$, the limit of $\frac{\dd{\cal G(0)}}{\dd z}$ satisfies that
\begin{align}\label{cal G'0}
	\frac{\dd \cal G(0)}{\dd z} \to \frac{-\frac{\dd a_1(0)}{\dd z}\Lambda^2 - (\frac{\dd a_2(0)}{\dd z} - 1)\id}{(a_{1}(0)\Lambda^2 + a_{2}(0)\id_p)^2}.
\end{align}
To find the derivatives of $a_1(z)$ and $a_2(z)$, we take the derivatives on both sides of the system of equations \eqref{selfomega_a}.
%\begin{align*}
%	\frac{\dd a_1(z)}{\dd z} + \frac{\dd a_2(z)}{\dd z} = -\frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{1}{\lambda_i^2 a_1 + a_2},
%	\frac{\dd a_1(z)}{\dd z} + \frac{1}{n_1 + n_2}\sum_{i=1}^p \frac{\lambda_i^2 (a_1'(z) a_2 - a_2'(z) a_1)}{(\lambda_i^2 a_1 + a_2)^2} = -\frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 a_1}{(\lambda_i^2 a_1 + a_2)^2}
%\end{align*}
Let $a_3 = - \frac{\dd a_1(0)}{\dd z}$ and $a_4 = - \frac{\dd a_2(0)}{\dd z}$.
%then taking implicit differentiation of equation \eqref{selfomega_a}
One can verify that $a_3$ and $a_4$ satisfy the self-consistent equations in \eqref{eq_a34extra} (details omitted).
Applying equation \eqref{cal G'0} to equation \eqref{calculate G'}, we obtain the asymptotic limit of the bias term.
%\begin{align}
%& n^2\bignorm{\Sigma_2^{1/2} \bigbrace{ (X^{(1)})^{\top}(X^{(1)}) + (X^{(2)})^{\top}(X^{(2)}) }^{-1} \Sigma_1^{1/2} w}^2 \nonumber\\
%&\approx  w^\top \Sigma_1^{1/2}\Sigma_2^{-1/2}V\frac{a_3\Lambda^2 +(1+ a_4)\id_p}{(a_{1}\Lambda^2 + a_{2}\id)^2}V^\top \Sigma_2^{-1/2}\Sigma_1^{1/2}w= w^{\top} \Pi_\bias w, \label{calculatePibias}
%\end{align}
%where in the last step we used $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ and $V \Lambda^2 V^\top=M^\top M$. This concludes equation \eqref{lem_cov_derv_eq}.

@@

Next, we introduce the bounded support assumption for a random matrix.
We say that a random matrix $Z \in\real^{n \times p}$ satisfies the {\it{bounded support condition}} with $Q$ or $Z$ has support $Q$ if
\begin{equation}
	\max_{1\le i \le n, 1 \le j \le p}\vert Z_{i j} \vert \prec Q. \label{eq_support}
\end{equation}
As shown in the example above, if the entries of $Z$ have finite moments up to any order, then $Z$ has bounded support $1$.
More generally, if the entries of $Z$ have finite $\varphi$-th moment, then using Markov's inequality and a simple union bound we get that %\HZ{I don't fully get the equation below}
\begin{align}
	\P\left(\max_{1\le i\le n, 1\le j \le p}|Z_{i  j}|\ge (\log n) n^{\frac{2}{\varphi}}\right) &\le \sum_{i=1}^n \sum_{j=1}^p \P\left(|Z_{i j}|\ge (\log n) n^{\frac{2}{\varphi}}\right)  \nonumber\\
	&\le \sum_{i=1}^n \sum_{j=1}^p \frac{C(\varphi)}{\left[(\log n) n^{\frac{2}{\varphi}}\right]^\varphi} = \OO((\log n)^{-\varphi}).\label{Ptrunc}
	\end{align}
In other words, $Z$ has bounded support $Q=n^{\frac{2}{\varphi}}$ with high probability.
%This leads to the exponent $\frac{\varphi-4}{2\varphi}$ in Fact \ref{lem_minv}. %This fact will be used in the proof of Corollary \ref{main_cor} below.
%The above discussion shows that the bounded support condition provides great flexibility in dealing with finite moment conditions.
%\begin{assumption}%\label{assm_big1}
%As in Section \todo{add reference}, we consider random matrices $X^{(1)}=Z^{(1)}\Sigma_1^{1/2}$ and $X^{(2)}=Z^{(2)}\Sigma_2^{1/2}$, where $Z^{(1)}$ and $Z^{(2)}$ are independent $n_1\times p$ and $n_2\times p$ random matrices with i.i.d. entries of zero mean and unit variance. Recall that $\rho_1= n_1/p$ and $\rho_2=n_2/p$. We assume that for a small constant $0<\tau<1$,
%\HZ{We have defined $\rho_1,\rho_2$ already in sec 2. Consider using ``Recall that $\rho_1 =, \rho_2 =$.''}
%\be\label{assm2}
%0\le \rho_1 \le \tau^{-1}, \quad 1+\tau \le \rho_{2} \le \tau^{-1}.
%\ee
%Recall that the singular value decomposition of $M = \Sig_1^{1/2} \Sig_2^{-1/2}$ is $U \Lambda V^{\top}$. We assume that
%\begin{equation}\label{assm32}
%\tau \le \lambda_p \le \lambda_1 \le \tau^{-1} .%, \quad \max\left\{\pi_A^{(n)}([0,\tau]), \pi_B^{(n)}([0,\tau])\right\} \le 1 - \tau .
%\end{equation}
%Finally, recalling the notations in equation \eqref{labelZ}, we assume that $Z^{(1)}$ and $Z^{(2)}$ satisfy
%\begin{equation}
%n^{-1/2}\max_{i\in \cal I_0,\mu \in \cal I_1}\vert Z^{(1)}_{\mu i}\vert \prec q,\quad n^{-1/2}\max_{i\in \cal I_0,\nu \in \cal I_2}\vert Z^{(2)}_{\nu i}\vert \prec q, \label{eq_support}
%\end{equation}
%for some deterministic parameter $q$ satisfying $ n^{-{1}/{2}} \leq q \leq n^{- \tau} $ for a small constant $\tau>0$.
%Here we recall the notations in \eqref{labelZ}.
 %and (\ref{assm2}). We assume that $T$ is an $M\times M$ deterministic diagonal matrix satisfying (\ref{simple_assumption}) and (\ref{assm3}).
%\end{assumption}
%\HZ{This is restating what's stated above again. Remove redundant statements (unless there's something critical missing in Sec 2).} [\cor FY: The setting in sec 1 is too vague for the purpose of this section in the sense that it is hard to find the exact properties we need. In this section, I want to state rigorous statements with a reference to an exact assumption. Moreover, the above assumption contains more.\nc]
